Evaluating the Performance and Energy Efficiency of COSMO-ART, a Coupled Numerical Weather Forecast and Chemical Transport Model

### 2014-04-19 paper_enahpc14/Sec2:COSMO-ART.tex
New version

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of $0.22\,^{\circ}$,  i.e.  50 km
in  both directions and  40 vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the  model initialization there is a  period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

The meteorological initial and  bounday conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at 6h temporal resolution. The  model setup incorporates 34 2-d and 45
3-d  fields  to be  written  out  every hour  and  is  devoid of  data
assimilation methods.

The  considered  COSMO-ART model  system  is  configured  with a  semi
Lagrangian horizontal advection scheme with tricubic interpolation and
selective filling  diffusion option in combination  with the dynamical
core  using Runge-Kutta  time  stepping \citep{COSMO-PartI-2011}.   It
also  makes use  of  the  Kinetic PreProcessor  solver  (KPP) for  the
resolution  of atmospheric  chemistry ordinary  differential equations
\citep{Damian-2002}.   Concerning the modelling  of wet  deposition in
aerosols, the  baseline has only indirect cloud  feedbacks but doesn't
include  in-cloud  scavenging  (rainout)  and  below-cloud  scavenging
(washout)  yet.   Amongst  physical  parameterisations,  precipitation
formation is performed by a two-moment cloud microphysics.

>>>>

Three-dimensional simulations are performed over large parts of Europe
and the Mediterranean~Sea for April 13th 2010 in order to benefit from
around  half  day  of  sun  exposure and  therefore  ensure  a  proper
activation of the chemistry cycle.  The calculation domain corresponds
to the CORDEX-EU-44 domain and is covered by a grid of $222\times 216$
points with  a horizontal resolution of $0.22\,^{\circ}$,  i.e.,  50~km
in  both directions and  40~vertical  layers.  These  24-hour forecast
simulations  are not  preceded  by  a spin-up  phase  to build-up  the
simulated gaseous and  aerosol concentrations.  This condition entails
that at the  model initialization there is a  period of adjustment for
the model  to reach its equilibrium  state and minimize  the effect of
initial conditions for gases and aerosols on the model predictions.

The meteorological initial and  boundary conditions are obtained by the
the ECMWF  global spectral model IFS  with an update  frequency of 3h.
Boundary data  for gas-phase species are taken  from IFS-MOZART output
at 6h temporal resolution. The  model setup incorporates 34~2-D and
45~3-D  fields  to be  written  out  every hour  and  is  devoid of  data
assimilation methods.

The  considered  COSMO-ART model  system  is  configured  with a  semi
Lagrangian horizontal advection scheme with tricubic interpolation and
selective filling  diffusion option in combination  with the dynamical
core  using Runge-Kutta  time  stepping \citep{COSMO-PartI-2011}.   It
also  makes use  of  the  Kinetic PreProcessor  solver  (KPP) for  the
resolution  of atmospheric  chemistry ordinary  differential equations
\citep{Damian-2002}.   Concerning the modelling  of wet  deposition in
aerosols, the  baseline has only indirect cloud  feedbacks but does not
include  in-cloud  scavenging  (rainout)  and  below-cloud  scavenging
(washout)  yet.   Amongst  physical  parameterisations,  precipitation
formation is performed by a two-moment cloud microphysics.

### 2014-04-19 paper_enahpc14/paper.tex
New version

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}
%
\smartqed
%
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{units}
\usepackage{color}
\usepackage{xspace}
%
\DeclareRobustCommand\IPCClongname{}
%
% please place your own definitions here and don't use \def but \newcommand{}{}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%
\begin{document}

\title{Evaluating the Performance  and Energy Efficiency of COSMO-ART,
  a Coupled Numerical Weather Forecast and Chemical Transport Model}
% \thanks{This work was supported by the EU Project FP7 318793 ``EXA2GREEN''}}
%\subtitle{Write here subtitle}

@@

\author{Joseph~Charles \and William~Sawyer \and Manuel~F.~Dolz \and
  Sandra~Catal\'an \and Cristiano~Malossi }

@@

\institute{J.~Charles, W.~Sawyer 
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland  
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch} 
        \and M.~F.~Dolz 
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}  
        \and S.~Catal\'an
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}
        \and C.~Malossi  
        \at IBM Zurich Research Laboratory 
        \\ CH-8000 Zurich, Switzerland
        \\ \email{ACM@zurich.ibm.com}
}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\RequirePackage{fix-cm}
%
\documentclass[twocolumn]{svjour3}
%
\smartqed
%
\usepackage{graphicx}
\usepackage[USenglish]{babel}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\usepackage[round, sort&compress, numbers]{natbib}
\usepackage{multirow}
\usepackage{graphicx}
\usepackage{units}
\usepackage{color}
\usepackage{xspace}
%\usepackage{cite}
%
\DeclareRobustCommand\IPCClongname{}
%
% please place your own definitions here and don't use \def but \newcommand{}{}
\newcommand{\pmlib}{\texttt{pmlib}\xspace}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%
\def\plusNoSpace{\texttt{+}}
\def\plus{\texttt{+}\xspace}
\def\minusNoSpace{\texttt{-}}
\def\minus{\texttt{-}\xspace}
\def\cpp{C{}\plusNoSpace\plusNoSpace\xspace}

\begin{document}

\title{Evaluating the Performance  and Energy Efficiency of COSMO-ART,
  a Coupled Numerical Weather Forecast and Chemical Transport Model\thanks{IBM, the IBM~logo, ibm.com, and Blue~Gene/Q are trademarks or registered trademarks of International Business Machines Corporation in the United~States, other countries, or both.}}
% \thanks{This work was supported by the EU Project FP7 318793 ``EXA2GREEN''}}
%\subtitle{Write here subtitle}

@@

\author{Joseph~Charles \and William~Sawyer \and Manuel~F.~Dolz \and
  Sandra~Catal\'an \and A.~Cristiano~I.~Malossi \and Costas~Bekas}

@@

\institute{J.~Charles, W.~Sawyer
	\at Swiss National Supercomputing Centre (CSCS)
	\\ CH-6900 Lugano, Switzerland
 	\\ E-mails:~{\{joseph.charles,william.sawyer\}@cscs.ch}
        \and M.~F.~Dolz
        \at Dept. of Informatics, University of Hamburg (UHAM)
	\\ DE-22527 Hamburg, Germany
        \\ \email{manuel.dolz@informatik.uni-hamburg.de}
        \and S.~Catal\'an
        \at Jaume I University of Castell\'on (UJI)
	\\ ES-12071 Castell\'on, Spain
        \\ \email{catalans@icc.uji.es}
        \and A.~C.~I.~Malossi, C.~Bekas
        \at IBM Research -- Zurich
        \\ CH-8803 R\"uschlikon, Switzerland
        \\ E-mails:~{\{acm,bek\}@zurich.ibm.com}
}

@@

\section{HPC Supercomputers}
\label{sec:5}
\input{Sec5:Supercomputers.tex}
